---
title: "Statistical Inference Course Project, Simulations"
author: "Jackson"
date: "November 23, 2015"
output: html_document
---


The exponential distribution can be simulationsulated in R with `rexp(n, lambda)` where
`lambda` $\lambda$ is the rate parameter. The mean of exponential distribution is 
$1/\lambda$ and the standard deviation is also $1/\lambda$. For this simulationsulation,
we set $\lambda=0.2$. In this simulations project, i will investigate the distribution of
averages of 40 numbers sampled from exponential distribution with $\lambda=0.2$.

There are 1000 simulationsulated averages of 40 exponentials for the tests.


```{r}
#load libraries 
library(ggplot2)

# create reproducability
set.seed(3)
# set contants
lambda <- 0.2
num_simulation <- 1000
sample_size <- 40
# test the result in num_simulations matrix
simulations <- matrix(rexp(num_simulation*sample_size, rate=lambda), num_simulation, sample_size)
# the mean of distribution of averages of exponentials
row_means <- rowMeans(simulations)
```

The distribution of sample means is as shown below.

```{r echo=FALSE}
# plot the histogram
hist(row_means, breaks=50, prob=TRUE,
     main="Distribution Of Averages Of Samples,
     drawn from exponential distribution with lambda=0.2",
     xlab="")
# density of the averages of samples
lines(density(row_means))
# theoretical center of distribution
abline(v=1/lambda, col="red")
# theoretical density of the averages of samples
xfit <- seq(min(row_means), max(row_means), length=100)
yfit <- dnorm(xfit, mean=1/lambda, sd=(1/lambda/sqrt(sample_size)))
lines(xfit, yfit, pch=22, col="red", lty=2)
# add legend
legend('topright', c("simulationsulation", "theoretical"), lty=c(1,2), col=c("black", "red"))
```

The distribution of sample means is at centered  `r mean(row_means)`
and the theoretical center of the distribution is $\lambda^{-1}$ = `r 1/lambda`.
The variance of sample means is `r var(row_means)` where the theoretical variance
of the distribution is $\sigma^2 / n = 1/(\lambda^2 n) = 1/(0.04 \times 40)$ =
`r 1/(0.04 * 40)`.

The averages of samples is follow normal distribution, there is because of the central limit theorem. The figure has showns the density computed using the histogram and thenormal density plotted with theoretical mean and variance values. 


The q-q plot has shown the distribution is approximately normality.

```{r echo=FALSE}
qqnorm(row_means); qqline(row_means)
```

Lastly, the author has evaluated the coverage of the confidence interval with
$1/\lambda = \bar{X} \pm 1.96 \frac{S}{\sqrt{n}}$

```{r echo=FALSE}
lambda_vals <- seq(4, 6, by=0.01)
coverage <- sapply(lambda_vals, function(lamb) {
    mu_hats <- rowMeans(matrix(rexp(sample_size*num_simulation, rate=0.2),
                               num_simulation, sample_size))
    ll <- mu_hats - qnorm(0.975) * sqrt(1/lambda**2/sample_size)
    ul <- mu_hats + qnorm(0.975) * sqrt(1/lambda**2/sample_size)
    mean(ll < lamb & ul > lamb)
})

library(ggplot2)
qplot(lambda_vals, coverage) + geom_hline(yintercept=0.95)
```

There are 95% confidence intervals for the rate parameter ($\lambda$) to be estimated
($\hat{\lambda}$) are
$\hat{\lambda}_{low} = \hat{\lambda}(1 - \frac{1.96}{\sqrt{n}})$ and
$\hat{\lambda}_{upp} = \hat{\lambda}(1 + \frac{1.96}{\sqrt{n}})$.
As shown from the plot above, for selection of $\hat{\lambda}$ around 5,
the average of the sample mean falls within the confidence interval at least 95% of the time.
Therefore the true rate, $\lambda$ is 5.

The Author has included the code at ...
